### 2015-08-29 vmcai-2016-tex/intro.tex
Started new folder for VMCAI submission.

In recent years symbolic execution has provided the basis for various
software testing and analysis techniques. Symbolic execution explores
the program by treating input values as symbolic, and for each
explored path, it computes a set of feasible constraints on the
symbolic inputs to create a \emph{path condition}. Path conditions
computed by symbolic execution characterize the observed program
execution behaviors and have been used as an enabling technology for
various applications, e.g., regression
analysis~\cite{backes:2012,Godefroid:SAS11,Person:FSE08,person:pldi2011,Ramos:2011,Yang:ISSTA12},
data structure repair~\cite{KhurshidETAL05RepairingStructurally},
dynamic discovery of
invariants~\cite{CsallnerETAL08DySy,Zhang:ISSTA14}, and
debugging~\cite{Ma:2011}.

The path conditions computed by symbolic execution have two important
features: 1) they characterize the set of feasible execution paths,
i.e., infeasible paths are omitted, and 2) each path condition encodes
the set of concrete program inputs that would cause execution to
follow the same control flow path. Because path conditions are encoded
as constraints over program inputs, the reasoning capabilities of
symbolic execution are limited by the underlying constraint
solver. Recent work has explored the strengths and limitations of
tools which leverage constraint solvers, e.g., concolic execution,
identifying symbolic dereferencing and updating operations as areas
for potential improvement~\cite{Chen:2013,Qu:2011}.  Precisely
modeling these operations, however, is challenging because it requires
formulating logical predicates over an input domain that contains a
potentially unbounded number of references.

The most direct approach to address these challenges is to approximate
the results of a symbolic deference by substituting in a valid
concrete value, as is done in dynamic symbolic execution
(\dsetxt{})~\cite{Godefroid:POPL07,Godefroid:2005,Sen:2005,Tillmann:2008}. These
approaches partition the universe of heaps in the symbolic
representation but often under-approximate the partitions or require
array theories in the constraint
solver~\cite{Chen:2013,Elkarablieh:2009}. A more effective approach is
to partition the reachable heaps by using value summary
representations, where multiple values for a reference may be
represented along a single execution
path~\cite{Dillig:2011,Elkarablieh:2009,Sen:2014,Torlak:2014,Xie:2005,Yorsh:2008}. Systems
using these representations perform analyses that are more complete
than standard~\dsetxt{} methods, but their associated dereferencing
operations require placing further restrictions on the input heap,
such as limiting the number of nodes~\cite{Elkarablieh:2009}, the
nature of aliasing relationships~\cite{Babic:2007,Xie:2005}, or the
the use of recursive data structures~\cite{Dillig:2011}.

An alternative approach which does not under-approximate the heap is
Generalized Symbolic Execution
(\gsetxt{})~\cite{Cadar:2008,KiasanKunit,GSE03,Rosner:2015}.  On symbolic
dereferencing operations,~\gsetxt{} partitions the input space between the
various possible aliasing configurations on-the-fly and also creates
non-deterministic choice points on the current path in order to
explore each heap configuration. This approach, however, is
inefficient because it initiates a new search path for every possible heap configuration.
The resulting path growth is exponential in the number of dereferencing operations.

The contribution of this work is a method for modeling and solving systems 
of constraints over a symbolic input heap.The
result of which, unlike~\gsetxt{}, is the ability to efficiently generate
and update \emph{symbolic heaps} to produce a single path constraint
that is sound and complete up to the point of execution. This is
accomplished by representing the symbolic heap with a field sensitive
points-to graph with edges annotated with disjunctive formulas
describing aliasing relationships. Where~\gsetxt{} branches the heap at every
de-reference, the symbolic heap in this work splits only at reference
comparison operations or field operations that may throw exceptions.

The correctness of the symbolic heap and its on-the-fly update rules
for symbolic execution is shown by proving the existence of a
bisimulation between symbolic heap states and~\gsetxt{} states. From that
result comes the soundness and completeness of the path constraint up
to the point of execution.  The efficiency of the representation is
established with an exploratory empirical evaluation within the Java
PathFinder framework over a classical set of symbolic execution
benchmarks. The evaluation demonstrates how the symbolic heap scales
compared to state-of-art~\gsetxt{} algorithms.

This new symbolic heap and update rules for symbolic execution is a
novel approach for de-referencing and updating heap allocated data
that brings together the best features of previous work while removing
the limitations of that same work. The symbolic heap has the following 
properties:
\begin{compactitem}
  \item constraints are neither over- or under-approximated;
  \item it does not restrict the set of heaps that can be defined;
  \item it does not require explicit construction of heap inputs;
  \item automatically computes path pre- and postconditions;
  \item generates path-coverage-optimal sets of test input heaps;
  \item it supports complex user-defined heap assertions 
\end{compactitem}

>>>>

In recent years symbolic execution has provided the basis for various
software testing and analysis techniques, e.g., regression
analysis~\cite{backes:2012,Godefroid:SAS11,Person:FSE08,person:pldi2011,Ramos:2011,Yang:ISSTA12},
data structure repair~\cite{KhurshidETAL05RepairingStructurally},
dynamic discovery of
invariants~\cite{CsallnerETAL08DySy,Zhang:ISSTA14}, and
debugging~\cite{Ma:2011}.

Recent work has explored the strengths and limitations of
tools which leverage constraint solvers, e.g., concolic execution,
identifying symbolic dereferencing and updating operations as areas
for potential improvement~\cite{Chen:2013,Qu:2011}.  Precisely
modeling these operations, however, is challenging because it requires
formulating logical predicates over an input domain that contains a
potentially unbounded number of references.

A promising approach to this problem has been to
to partition the reachable heaps by using \emph{symbolic heap}
representations, where multiple values for a reference may be
represented along a single execution
path~\cite{Dillig:2011,Elkarablieh:2009,Sen:2014,Torlak:2014,Xie:2005,Yorsh:2008}. However, it remains an open question whether or not these methods can be used to exactly model all possible program behaviors in the case of an arbitrary input.

The contribution of this work is to build on these previous efforts, by demonstrating how to dereference pointers in an unknown symbolic input heap, thereby proving that value-summary representations are sound and complete with respect to properties provable by symbolic execution. A proof-of-concept implementation also demonstrates the efficiency of value-summary representations, as compared to the previous best-known exact analysis method.

### 2015-08-29 vmcai-2016-tex/summary-heap.tex
Started new folder for VMCAI submission.

In symbolic execution, the program state is represented symbolically, as a function of the program’s inputs. That is to say, the values stored in the memory locations accessed by the program are represented as expressions where the only free variables are the program input parameters. For programs accepting references as input parameters, the symbolic input heap exists as an additional, implicit input parameter. Symbolic reference values are expressions formed entirely in terms of the input reference parameters and the symbolic input heap.

@@

%\sjp{Up to now, no discussion of lazy initialization - what it is or
%  how it is being used.}  \nsr{A key insight to the summary heap is
%  that although many auxiliary or stack literals are created during
%  symbolic execution with lazy initialization, only input heap
%  references appear in the constraints expressing potential aliasing
%  relationships.}
A solver is used to reason over the aliasing relationships, and a
satisfying assignment uniquely identifies the set of aliasing
relationships that are representative of a concrete heap and its
shape. For example, if the solver assigns input heap variables
$\cfgnt{r}$ and $\cfgnt{r}^\prime$ such that $\cfgnt{r} =
\cfgnt{r}^\prime$ is true, then there exists some location $\cfgnt{l}$
and constraints $\phi$ and $\phi^\prime$ such that (i)
$\lp\phi\ \cfgnt{l}\rp \in \cfgnt{L}(\cfgnt{r})$; (ii)
$\lp\phi^\prime\ \cfgnt{l}\rp \in \cfgnt{L}(\cfgnt{r}^\prime)$; and
(iii) $\phi$ and $\phi^\prime$ hold in the aliasing assignments from
the solver.  In this way, the graph groups heaps with the conditions
under which those heaps exist so look-ups are non-recursive and
updates do not modify the path condition or require rewrites on
constraints.

@@

In this work, symbolic execution semantics are is defined using
a small-step operational semantics similar to the one presented 
in~\cite{saints-MS}. The small-step semantics,
defined as a syntactic machine using a CESK architecture~\cite{Felleisen:1992},
are modified to perform symbolic execution on a symbolic heap. The
surface (input) syntax and machine (state) syntax are shown
in~\figref{fig:syntax}. Terminals are in bold face while non-terminals
are italicized. Ellipses indicate zero or more repetitions. Tuples
omit the commas for compactness.

>>>>

@@

@@

Symbolic execution semantics are is defined using
a small-step operational semantics similar to the one presented 
in~\cite{saints-MS}. The small-step semantics,
defined as a syntactic machine using a CESK architecture~\cite{Felleisen:1992},
are modified to perform symbolic execution on a symbolic heap. The
surface (input) syntax and machine (state) syntax are shown
in~\figref{fig:syntax}. Terminals are in bold face while non-terminals
are italicized. Ellipses indicate zero or more repetitions. Tuples
omit the commas for compactness.
